Title

### 2025-04-14 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
NACFC joint model update

```{r setup}
#| include: false
library(survival)
library(nlme)
library(JM)
library(dplyr)
library(tidyr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

@@

# Make survival datasets for registry and OGTT diagnoses
ogtt <- annual %>%

@@

    eDWID, ReviewYear, Age_YrEnd, A_hgba1c, all_of(predictors), cfrd_ogtt

@@

surv_ogtt <- annual %>%
  filter(Age_YrEnd <= cfrd_ogtt_age) %>%

@@

  select(eDWID, ReviewYear, Age_YrEnd, all_of(predictors), cfrd_ogtt) %>%

@@

ogtt_include <- intersect(ogtt$eDWID, surv_ogtt$eDWID)

@@

@@

```{r}
# Formulas
cox_f <- as.formula(paste0(
  "Surv(time = Age_YrEnd, event = cfrd_ogtt)~",
  paste0(predictors, collapse = "+")
))
lme_f <- as.formula(paste0(
  "A_hgba1c~Age_YrEnd+",
  paste0(predictors, collapse = "+")
))
# Fit Cox model first, and then pull the clusters
cox_fit <- coxph(cox_f,
  cluster = eDWID, x = T, model = T,
  data = surv_ogtt[surv_ogtt$eDWID %in% ogtt_include, ]
)
lme_fit_a1c <- lme(lme_f,
  random = ~ Age_YrEnd | eDWID, na.action = na.omit,
  data = ogtt[ogtt$eDWID %in% ogtt_include, ]
)
# Combine
joint_fit <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd")
```

@@

>>>>

```{r setup}
#| include: false
library(survival)
library(nlme)
library(JM)
library(dplyr)
library(tidyr)
library(arsenal)
library(knitr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

annual$Gender <- factor(annual$Gender, levels = c("M", "F"))

@@

# Make datasets
registry <- annual %>%
  filter(Age_YrEnd <= cfrd_registry_age) %>%

@@

    eDWID, ReviewYear, Age_YrEnd, A_hgba1c, all_of(predictors), cfrd_registry

@@

surv_reg <- annual %>%
  filter(Age_YrEnd <= cfrd_registry_age) %>%

@@

  select(eDWID, ReviewYear, Age_YrEnd, all_of(predictors), cfrd_registry) %>%

@@

reg_include <- intersect(registry$eDWID, surv_reg$eDWID)

@@

# Demographics

```{r,results='asis'}
t <- registry[registry$eDWID %in% reg_include, ] %>%
  group_by(eDWID) %>%
  summarise(n())

t1 <- tableby(cfrd_registry ~ Age_YrEnd + Gender + Race + Hispanicrace + A_IsOnEnzymes + NumPulmExacerbation + A_FEV1_pct_predicted + bmi, data = surv_reg[surv_reg$eDWID %in% reg_include, ])
summary(t1)
```

@@

```{r}
# Formulas
cox_f <- as.formula(paste0(
  "Surv(time = Age_YrEnd, event = cfrd_registry)~",
  paste0(predictors, collapse = "+")
))
lme_f <- as.formula(paste0(
  "A_hgba1c~Age_YrEnd+",
  paste0(predictors, collapse = "+")
))
# Fit Cox model first, and then pull the clusters
cox_fit <- coxph(cox_f,
  cluster = eDWID, x = T, model = T,
  data = surv_reg[surv_reg$eDWID %in% reg_include, ]
)
lme_fit_a1c <- lme(lme_f,
  random = ~ Age_YrEnd | eDWID, na.action = na.omit,
  data = registry[registry$eDWID %in% reg_include, ]
)
# Combine
joint_fit_value <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd", parameterization = "value")
# joint_fit_slope <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd",parameterization ="slope")
# joint_fit_both <- jointModel(lme_fit_a1c, cox_fit, timeVar = "Age_YrEnd",parameterization ="both")
s <- summary(joint_fit_value)
long <- data.frame(s$`CoefTable-Long`)
long <- long[, c(1, 4)]
long$p.value <- format.pval(long$p.value, eps = 0.001, digits = 3)
event <- data.frame(s$`CoefTable-Event`)
event <- event[, c(1, 4)]
event$Value <- exp(event$Value)
rownames(event)[nrow(event)] <- "shape"
event$p.value <- format.pval(event$p.value, eps = 0.001, digits = 3)
# summary(joint_fit_slope)
# summary(joint_fit_both)
kable(long, caption = "Longitudinal Process", digits = 3)
kable(event, caption = "Event Process", digits = 3)
```

# To do

@@

https://erandrinopoulou.github.io/EducationalCorner_JMpred/

### 2025-04-14 Quarto Template.qmd
NACFC joint model update

```{r data cleaning}
#| message: false
# Import from REDCap
unlockREDCap(c(rcon = ""),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
df <- exportRecordsTyped(rcon)
```

>>>>

```{r data cleaning}
# Import from REDCap
unlockREDCap(c(rcon = ""),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
df <- exportRecordsTyped(rcon)
```
